Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation

### 2014-03-05 writedown/npt-neq/reply/reply.tex
2nd round

We have employed equilibrium simulations to obtain the non-equilibrium
responses within the framework of linear responses theory (Green-Kubo relation).
With this approach we studied the conformational shift of alanine dipeptide under the electric field
(EF). As response function we have considered the time-dependent probability of
left-handed $\alpha$-helix (Conformation $C$ in the old manuscript. The notation has been changed to $\alpha_L$ in the updated manuscript.) This is defined as
\begin{align}\label{eqn:tmp1}
  P_C(t) = \langle\chi_{_C}(t)\rangle = \langle \chi_{_C} \rangle_0 -
  \beta \int_0^t ds\; M_e(t - s)\langle j(0)\cdot \chi_{_C}(t) \rangle_0
\end{align}
where $\chi_{_C}$ is the characteristic function of set $C$ that takes
the value of 1 for $(\phi,\psi)\in C$, and takes value 0
otherwise. $M_e(t)$ is the magnitude of the EF. In the case of
constant EF, $M_e(t) = t/t_{warm}$ for $0\leq t<t_{warm}$, and
$M_e(t) = 1$ for $t\geq t_{warm}$.
$j$ is the dissipative flux that is defined by
\begin{align}
  j = - \sum_{i=1}^N E_\infty q_i v_{i,x},
\end{align}
where $q_i$ is the partial charge of the $i$th atom, and $v_{i,x}$ is
the $x$ component of the velocity of the $i$th atom.  The notation
$\langle\cdot\rangle_0$ denotes the equilibrium ensemble average.
 We then plot the correlation function
$\langle j(0)\cdot \chi_{_C}(t) \rangle_0$ against correlation time $t$
in Fig.~\ref{fig:tmp1} of this reply. The value of the function is estimated from
two independent equilibrium simulations, and the trajectory of each is
of 1~$\mu$s. The total length of the equilibrium
trajectories is the same as the non-equilibrium simulation (2000
trajectories, each 1000~ps long).

We find that the statistical error is even larger than the value of
correlation function.  Therefore, the
overwhelming statistical uncertainty does not allow us to use the equilibrium approach 
in Eq.~\eqref{eqn:tmp1} to calculate the
non-equilibrium averages. 
% , and we actually do not know when the
% corrlation function vanish .
Since the total computational effort of
the equilibrium simulation is the same as the non-equilibrium
simulation, we conclude that non-equilibrium simulation is more
accurate and efficient than the conventional equilibrium approach  at the same computational cost.
It must be noticed that a similar
observation has already been reported by Ciccotti and Jacucci, (Phys.~Rev.~Lett.~35, 789, 1975). We report the description above in the appendix of the manuscript.
\\
\begin{figure}
  \centering
  \includegraphics{figs/fig-corr-meta.eps}
  \caption{The correlation function $\langle j(0)\cdot \chi_{_C}(t)
    \rangle_0$ against time $t$. The error bars in the figure denote
    the statistical uncertainty at 95\% confidence level. }
  \label{fig:tmp1}
\end{figure}

@@

The question on whether our non-equilibrium branching trajectories are
long enough so that the results are fully converged is a very subtle
question. There may exists an intrinsic time-scale longer than our simulation time, however this is obvious and implicitly imply that the validity of our results is restricted to the time scale simulated. 
\\

@@

The periods 10, 40~ps are roughly the same as the first two shortest
time scales of probability fluxes observed in the constant EF case,
see Table~I of the manuscript. We tried to simulation with period of
500~ps, which is the same as the longest timescales of the constant EF
case, however, the required simulation time is prohibitive at this stage. Therefore, we made a reasonable compromise and considered only
a period of 200~ps. The typical frequency of the microwave
irradiation is 2450~MHz, which corresponds to period of 400~ps. It is
the same order of magnitude as the $T_P=200$~ps studied by us.
We have added a comment along these lines in the revised manuscript.
\\

@@

Yes, the definition of the probability fluxes is similar to the
kinetic rate (we use joint probabilities in (6) rather than
conditional probabilities). However, we want to stress out that, in general,
the probability fluxes calculated in the manuscript should not be
interpreted as a kinetic rate, and one cannot replace the
non-equilibrium molecular dynamics by a kinetic analysis. The main
reasons are:
Firstly, for a non-equilibrium system, the probability fluxes are
time-dependent, therefore, the lag-time $\Delta t$ in the definition (Eq.~(6))
should be small compared with the typical time-scale of conformation change.\\
Secondly, for a small lag time $\Delta t$ (1~ps in the manuscript), the conformational
dynamics of biomolecules is generally not
Markovian (see e.g.~C.~Sch\"utte et.~al.~JCP 2011 and J.-H.~Prinz et.~al.~JCP 2011).
Since the basic assumption -- Markovianity -- of the kinetic analysis
may not be satisfied, a direct
application of the kinetic analysis in the non-equilibrium cases is questionable. Although one may
gain useful information from the kinetic analysis, the reliability
and the effectiveness of this approach should be carefully checked case
by case.  Since the main goal of our research is to propose a
non-equilibrium molecular dynamics simulation method, the numerical
results are not compared with a kinetic analysis.\\

A comment along this direction has been added to the end of section III.B.
\\

@@

The frequencies investigated in our manuscript span from 5~GHz to 100~GHz,
the largest of which is still one order of magnitude smaller than those normally used in  terahertz spectroscopy. Therefore, we find our simulation results, at this stage, cannot be directly compared with the terahertz experiments.
{To make fully aware the readers about the current limitation of our study, with respect to experimental connections, we have added the comment above in the manuscript.}
\\

@@

This is indeed a very interesting point.
Although the network presented Fig.11 is similar to those
discovered in the mentioned equilibrium studies,
regarding the map of the steady conformations and connections among them,
the physical meaning is instead quite different in our study:
The system is under the nonequilibrium conditions, and
the fluxes are explicitly induced by a physically realistic control, i.e.~EF, although the averaged population of the steady conformations do not change. In contrast,
such probability fluxes do not exist, as a result of an explicit physical action, in the equilibrium case.
A comment along such direction has been added to the
manuscript.\\

It must be also added that the transition fluxes can be used to calculate the free energy only in
the equilibrium case. In the non-equilibrium case, especially for the
oscillatory EF case, the conformational kinetics of the system is generally
not Markovian. Therefore, the standard kinetic analysis cannot be used
here.

>>>>

We have employed equilibrium simulations to obtain the non-equilibrium
responses within the framework of linear responses theory (Green-Kubo relation).
With this approach we studied the conformational shift of alanine dipeptide under the electric field
(EF). As response function we have considered the time-dependent probability of
left-handed $\alpha$-helix (Conformation $C$ in the old manuscript. The notation has been changed to $\alpha_L$ in the updated manuscript.) This is defined as
\begin{align}\label{eqn:tmp1}
  P_C(t) = \langle\chi_{_C}(t)\rangle = \langle \chi_{_C} \rangle_0 -
  \beta \int_0^t ds\; M_e(t - s)\langle j(0)\cdot \chi_{_C}(t) \rangle_0
\end{align}
where $\chi_{_C}$ is the characteristic function of set $C$ that takes
the value of 1 for $(\phi,\psi)\in C$, and takes value 0
otherwise. $M_e(t)$ is the magnitude of the EF. In the case of
constant EF, $M_e(t) = t/t_{warm}$ for $0\leq t<t_{warm}$, and
$M_e(t) = 1$ for $t\geq t_{warm}$.
$j$ is the dissipative flux that is defined by
\begin{align}
  j = - \sum_{i=1}^N E_\infty q_i v_{i,x},
\end{align}
where $q_i$ is the partial charge of the $i$th atom, and $v_{i,x}$ is
the $x$ component of the velocity of the $i$th atom.  The notation
$\langle\cdot\rangle_0$ denotes the equilibrium ensemble average.
 We then plot the correlation function
$\langle j(0)\cdot \chi_{_C}(t) \rangle_0$ against correlation time $t$
in Fig.~\ref{fig:tmp1}~(a) of this reply. The value of the function is estimated from
two independent equilibrium simulations, and the trajectory of each is
of 1~$\mu$s. The total length of the equilibrium
trajectories is the same as the non-equilibrium simulation (2000
trajectories, each 1000~ps long).
We find that the statistical error is even larger than the value of
correlation function.
In Fig.~\ref{fig:tmp1}~(b),
  the probability of conformation $C$ calculated from
  the linear response theory is compared with that
  computed from non-equilibrium simulation.
  The linear response result follows the non-equilibrium result only in the very first
  15~ps, then it diverges. Since it is meaningless for a probability
  being larger than 1, the linear response result is qualitatively wrong.
Therefore, the
overwhelming statistical uncertainty does not suggest to use the equilibrium approach 
in Eq.~\eqref{eqn:tmp1} to calculate the
non-equilibrium averages. 
% , and we actually do not know when the
% corrlation function vanish .
Since the total computational effort of
the equilibrium simulation is the same as the non-equilibrium
simulation, we conclude that non-equilibrium simulation is more
accurate and efficient than the conventional equilibrium approach  at the same computational cost.
It must be noticed that a similar
observation has already been reported by Ciccotti and Jacucci, (Phys.~Rev.~Lett.~35, 789, 1975). We report the description above in the appendix of the manuscript.
\\
\begin{figure}
  \centering
  \includegraphics{figs/fig-corr-meta.eps}
  \includegraphics{figs/fig-corr-meta-c.eps}
  \caption{
    The numerical result of the linear response theory.
    (a): The correlation function $\langle j(0)\cdot \chi_{C}(t)
    \rangle_0$ against time $t$.
    The error bars in the figure denote
    the statistical uncertainty at 95\% confidence level. 
    (b): The probability of conformation $C$ computed from the
    linear response theory versus that
    computed from non-equilibrium simulation (D-NEMD).}
  \label{fig:tmp1}
\end{figure}

@@

The question on whether our non-equilibrium branching trajectories are
long enough so that the results are fully converged is an always present problem. There may exists an intrinsic time-scale longer than our simulation time, however this is obvious and implicitly imply that the validity of our results is restricted to the time scale simulated. 
\\

@@

The periods 10, 40~ps are roughly the same as the first two shortest
time scales of probability fluxes observed in the constant EF case,
see Table~I of the manuscript. We tried to do a simulation with period of
500~ps, which is the same as the longest timescales of the constant EF
case, however, the required simulation time is prohibitive at this stage. Therefore, we made a reasonable compromise and considered only
a period of 200~ps. The typical frequency of the microwave
irradiation is 2450~MHz, which corresponds to period of 400~ps. It is
the same order of magnitude as the $T_P=200$~ps studied by us.
We have added a comment along these lines in the revised manuscript.
\\

@@

Yes, the definition of the probability fluxes is similar to the
kinetic rate (we use joint probabilities in (6) rather than
conditional probabilities).
However, we prefer to stick to our definition which
is mathematically simpler and univocal.
A comment along this direction has been added to the end of section III.B.
\\
% However, we want to stress out that, in general,
% the probability fluxes calculated in the manuscript should not be
% interpreted as a kinetic rate, and one cannot replace the
% non-equilibrium molecular dynamics by a kinetic analysis. The main
% reasons are:
% Firstly, for a non-equilibrium system, the probability fluxes are
% time-dependent, therefore, the lag-time $\Delta t$ in the definition (Eq.~(6))
% should be small compared with the typical time-scale of conformation change.\\
% Secondly, for a small lag time $\Delta t$ (1~ps in the manuscript), the conformational
% dynamics of biomolecules is generally not
% Markovian (see e.g.~C.~Sch\"utte et.~al.~JCP 2011 and J.-H.~Prinz et.~al.~JCP 2011).
% Since the basic assumption -- Markovianity -- of the kinetic analysis
% may not be satisfied, a direct
% application of the kinetic analysis in the non-equilibrium cases is questionable. Although one may
% gain useful information from the kinetic analysis, the reliability
% and the effectiveness of this approach should be carefully checked case
% by case.  Since the main goal of our research is to propose a
% non-equilibrium molecular dynamics simulation method, the numerical
% results are not compared with a kinetic analysis.\\

@@

The frequencies investigated in our manuscript span from 5~GHz to 100~GHz,
the largest of which is still one order of magnitude smaller than those normally used in  terahertz spectroscopy. Therefore, we find that our simulation results, at this stage, cannot be directly compared with the terahertz experiments.
{To make fully aware the readers about the current limitation of our study, with respect to experimental connections, we have added the comment above and the experimental references in the manuscript.}
\\

@@

This is indeed a very interesting point.
Although the network presented in Fig.11 is similar to those
discovered in the mentioned equilibrium studies,
regarding the map of the steady conformations and connections among them,
the physical meaning is instead quite different in our study:
The system is under nonequilibrium conditions, and
the fluxes are explicitly induced by a physically realistic control, i.e.~EF, although the averaged population of the steady conformations do not change. In contrast,
such probability fluxes do not exist, as a result of an explicit physical action, in the equilibrium case.
A comment along such direction has been added to the
manuscript.\\

% It must be also added that the transition fluxes can be used to calculate the free energy only in
% the equilibrium case. In the non-equilibrium case, especially for the
% oscillatory EF case, the conformational kinetics of the system is generally
% not Markovian. Therefore, the standard kinetic analysis cannot be used
% here.
